Self-assembly of colloid-cholesteric composites: A route to switchable optical materials

### 2013-10-03 papers/2012-sandwich/support.tex
Updated supplement
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@2172 40876441-3496-0410-a613-a2fca35b2ae2

In this work, the chirality and reduced temperature are always set to lie
within the blue phase I region of the equilibrium phase diagram. EXCEPT?

In addition to the fluid free density density $f(Q_{\alpha\beta})$, a surface
free energy density is also present to represent the energetic cost of
anchoring at a solid surface. If the preferred order parameter tensor
at the solid surface is $Q^0_{\alpha\beta}$, then the surface free energy
density (per unit area) is
\begin{equation}
f_s(Q_{\alpha\beta}, Q^0_{\alpha\beta})
= {\textstyle \frac{1}{2}}W(Q_{\alpha\beta} - Q^0_{\alpha\beta})^2,
\end{equation}
where $W$ is a constant determining the strength of the anchoring.
The determination of $Q^0_{\alpha\beta}$ for normal and planar anchoring is
discussed below. For a colloid of radius $a$, the strength of the surface
anchoring compared which the bulk fluid elastic constant may be quantified
by the dimensionless parameter $Wa/K$. For small values of this parameter,
the presence of a colloid surface should have little impact on the local
fluid LC ordering.

The effect of an electric field can be modelled by including the following
term in the bulk free energy density:
\begin{equation}
-\frac{\epsilon_a}{12\pi} E_{\alpha}Q_{\alpha\beta}E_{\beta},
\end{equation} 
where $E_{\alpha}$ are the components of the electric field, and 
$\epsilon_a$ (here assumed positive) is the dielectric anisotropy of the liquid
crystal. The strength of the electric field is quantified via
a further dimensionless number, 
\begin{equation}
{\cal E}^2 = \frac{27 \epsilon_a}{32 \pi A_0 \gamma} E_\alpha E_\alpha.
\end{equation}

@@

In all simulations the colloidal (hard sphere) radius is $a = 2.3$. To
prevent colloids overlapping at the hard sphere radius, an additional
short range soft potential is included. This takes the form of
$V(h) = \epsilon (\sigma/h)$ where $h$ is the surface to surface separation.
The parameters are $\epsilon = 0.0004$ and $\sigma = 0.1$ in simulation
units. Both potential and resulting force are smoothly matched to zero at
a cut-off distance of $h_c = 0.25$ simulation units.

@@

The parameters for the free energy are chosen to be representative
of blue phase I: chirality $\kappa \sim 0.7$ and reduced temperature
$\tau = -0.25$.
The full free energy parameters are shown in Table~\ref{table:params}.
The order parameter tensor $Q_{\alpha\beta}$ for bulk blue phase I is
initialised from an approximation in high chirality limit
\cite{blue1,oliver1}. Systems of 128$^3$ lattice sites are
used with a pitch length of $p = 32\sqrt{2}$, which accommodates 4$^3$
unit cells with fully periodic boundary conditions.
Simulations at different solid volume fractions (1\% and 4\%), and different 
surface anchoring strengths ($Wa/K = 0.23$ and $Wa/K = 2.3$) representing
``weak'' and ``strong'' anchoring are run for two million simulation
time steps.
The initial distribution of the colloid positions is set at random
for the required solid volume fraction; the colloids are initially
at rest.

To generate a disordered network of disclination lines, a ``quench''
is performed in which the order parameter is initialised via a
locally chosen random director, and
Eq.~\ref{eq:surf_q} employed with a small amplitude $q^0 = 10^{-7}$ to
provide an order parameter.
We avoid the mean-field spinodal point at $\gamma = 3.0$:
the parameters are the same as for the initial simulations with
pitch $p = 32\sqrt{2}$. The chirality and reduced temperature (see
Table~\ref{table:params}) remain appropriate for blue phase I.
Colloids are added as before and the simulations run for 2 million
simulation steps.

@@

Here, a narrow sandwich of fluid is placed between flat walls in
perpendicular to the narrow coordinate direction, and with periodic
boundaries in the other two directions. The system size used in all
cases is 256$^2 \times$56 lattice sites. Each simulation is a ``quench''
in a similar manner to that used in the bulk: the order parameter is
initialised randomly. The free energy parameters (see Table~\ref{table:params})
are again appropriate for (equilibrium) blue phase I
(chirality $\kappa = 0.8$ and
reduced temperature $\tau = -0.25$). The cholesteric pitch is set
to be $p = 64$, providing a slight better resolution than the bulk
simulations.

The surface anchoring for colloids is as before: always normal, but with
$Wa/K = 0.23$ and $Wa/K = 2.3$. The colloid solid volume fractions are
1\% and 4\%. Wall anchoring is either normal or planar with $W/K = 1$
in simulation units. All simulations are run for two million simulation
time steps.

@@

OLIVER PLEASE ADD short description of protocol here. AND the
parameters in the table.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
         & $A_0$ & $\gamma$ & $K$ & $p$ & $\kappa$ & $\tau$ & $Wa/K$\\
\hline 
Bulk BPI & 0.01 & 3.086 & 0.007061 & 32$\sqrt{2}$ & 0.6902 & -0.2500 &
0.23, 2.3\\
\hline
Bulk Quench & 0.01 & 3.086 & 0.007061 & 32$\sqrt{2}$ & 0.6902 & -0.2500 &
0.23, 2.3 \\
\hline
Confined & 0.01 & 3.086 & 0.01897 & 64 & 0.8 & -0.2500 & 0.23, 2.3\\
\hline
External field & a0& gamma & K & p & kappa & tau & Wa/K\\ 
\hline
\end{tabular}
\end{center}
\caption{Free energy parameters used in the simulations. Bulk BPI
correspond to Fig.~1 (A-D), bulk quench to Fig.~1 (E--H); confined
geometries are shown in Fig.~2; and the external field case is
relevant for Fig.~3.}
\label{table:params}
\end{table}

@@

To get from simulation to physical units, we need to calibrate scales of
length, energy and time in the simulations. For similar mappings see,
e.g.~\cite{denniston2}.

The length scale can be set by mapping our half pitch to a realistic blue
phase unit cell size, which are in the 100-500 nm range \cite{blue1}. 
For bulk simulations (Fig.~1 and Fig.~3 in the main text), a suitable choice
is one where one simulation unit (lattice site) corresponds to 10~nm. Therefore
the colloidal size in Figs.~1 and~3 corresponds to $\sim$ 50~nm.

To obtain an energy scale, we choose $A_0 \simeq 10^6$~Pa, which is 
reasonable from Ref.~\cite{blue1} (this is also the choice in
Ref.~\cite{oliver2}). This choice leads to a simulation unit of
free energy (or stress) equal to $10^{8}$~Pa in Figs.~1 and~3.
From the energy and length scales one can map elastic constants to
physical units: for instance, the simulation in Fig.~1 corresponds to
a liquid crystal with splay, bend and twist (Frank) elastic constants
equal to 35 pN (or equivalently $K=$70~pN). That is,
\begin{equation}
\frac{a}{b^2} = \frac{27}{2 A_0\gamma} \sim 2-5 \times 10^{-6}
{\rm J}^{-1}\, {\rm m}^{3}.
\end{equation}
WHAT ARE a/b???

For the timescale calibration we use the formula (see e.g.~\cite{denniston})
\begin{equation}
\gamma_1=\frac{2q^2}{\Gamma}
\end{equation}
which relates the rotational viscosity $\gamma_1$, to the ordering strength
$q$ and the order parameter mobility $\Gamma$.
In our simulations, we choose $\Gamma = 0.5$, while the value of $\gamma$
used in Fig.~1 leads to $q=1/2$.   
For real liquid crystalline materials, $\gamma_1$ usually lies in range
$10^{-2}-1$ Pa s~\cite{deGennes}; we choose for definiteness
$\gamma_1 = 1$~Pa~s (equivalently 10 poise). Given the previous mapping for
free energy (or stress) units, this leads to a simulation unit of time equal
to $10^{-8}$~s. The total simulated time in Figs. 1-3 is therefore in the
millisecond range.

To calibrate the electric field strength 
we equate the value of the dimensionless parameter ${\cal E}$ in our 
simulations with that of a hypothetical experiment.
Given the previous value for $A_0$, and a dielectric anisotropy of 10
(or equivalently $\epsilon_a=10\epsilon_0$, with $\epsilon_0$ the
dielectric permittivity of vacuum), we then obtain that an electric field
of 5~V/$\mu$m corresponds to a value of ${\cal E}\sim$0.01.

Given these mappings, all quantities can be easily transformed from
simulation to physical units and vice versa. For example Figs.~1 and~2
use an isotropic viscosity $\eta \simeq$~0.01 and~0.1 in simulation units,
which translate into $0.01-0.1$~Pa~s respectively. This is sensible (if 
slightly low) for a molecular nematogen in the isotropic phase. 
The effective viscosity in ordered phases is higher, but this is ensured
in our model by the coupling to the order parameter.
We note that the density is set to unity in our simulations: this
corresponds to a fluid density which is about a thousand times larger
than in experiments. This however makes no difference in practice as
the Reynolds number $= \rho V\Lambda/\eta$ (where $V$ is a typical
velocity, in our simulations around $10^{-5}$ in simulation units) 
remains small enough~\cite{codef}.

To summarize, our simulations represent BP-forming materials, with the
interpretation of the simulation units for length, time, and energy
density being close to 10~nm, 10~ns, and 100~MPa respectively.

@@

\bibitem{blue1}
D.C. Wright and N.D. Mermin,
%  Crystalline liquids -- the blue phases.
{\it Rev. Mod. Phys.} {\bf 61,} 385-432 (1989).

%\bibitem{marendu1}
%D. Marenduzzo, E.  Orlandini, M. E. Cates, J. M. Yeomans,
%{\rm Steady-state hydrodynamic instabilities of active liquid crystals:
%Hybrid lattice Boltzmann simulations.}
%{\it Phys. Rev. E} {\bf 76}, 031921 (2007).

%\bibitem{marendu2}
%O. Henrich, D. Marenduzzo, K. Stratford, M. E. Cates, 
%{\rm Domain growth in cholesteric blue phases:
%hybrid lattice Boltzmann simulations}.
%{\it Comput. Math. Appl.}, doi:10.1016/j.camwa.2009.08.047[dx.doi.org] (2009).

%\bibitem{marendu3}
%M. E. Cates, O. Henrich, D. Marenduzzo, K. Stratford,
%{\rm  Lattice Boltzmann simulations of liquid crystalline fluids:
%active gels and blue phases.}
%{\it Soft Matter} {\bf 5}, 3791-3800 (2009).

\bibitem{oliver1}
O. Henrich, D. Marenduzzo, K. Stratford, and M.E. Cates,
%Thermodynamics of blue phases in electric fields
\textit{Phys. Rev. E}, \textbf{81} 031706 (2010).

\bibitem{beris}
A. N. Beris, B. J. Edwards, 
{\it Thermodynamics of Flowing Systems with Internal Microstructure}
(Oxford Univ. Press, New York, 1994).

\bibitem{deGennes}
P.-G. de Gennes, J. Prost,
{\it The Physics of Liquid Crystals} (Oxford Univ. Press, New York, 1993).

\bibitem{ladd94}
A.J.C. Ladd,
\textit{J. Fluid Mech.} \textbf{271}, 285 (1994); \textit{ibid.} \textbf{271}
311 (1994).

\bibitem{nguyen2002}
N.-Q. Nguyen and A.J.C. Ladd,
\textit{Phys. Rev. E} \textbf{66} 046708 (2002).

\bibitem{denniston}
C. Denniston, E. Orlandini, J. M. Yeomans, 
\textit{Phys. Rev. E} \textbf{63}, 056702 (2001).

\bibitem{denniston2}
C. Denniston, D. Marenduzzo, E. Orlandini, J. M.  Yeomans, 
\textit{Philos. Trans. R. Soc. London, Ser. A} \textbf{362}, 1745 (2004).

\bibitem{fournier2005}
J.-B. Fournier and P. Galatola,
\textit{Europhys. Lett.} \textbf{72}, 403 (2005).

\bibitem{oliver2}
O. Henrich, K. Stratford, D. Marenduzzo, M. E. Cates, 
\textit{Proc. Acad. Sci. USA} \textbf{107}, 1312 (2010).

\bibitem{codef} M. E. Cates,  et al., 
\textit{J. Phys. Condens. Matter} \textbf{16},  3903 (2004).

@@

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.8]{doubletwist.pdf}
\end{center}
\caption{
(A) Schematic representation of the director field
in a cholesteric (left) and within a double twist cylinder (right).
(B) Snapshot of the disclination network in an equilibrated 
blue phase I structure. (C) Same as (B) for an amorphous blue phase III
network.}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.35]{support-fig2.pdf}
\end{center}
\caption{As for main manuscript Fig.~1 A--D
(with liquid crystal order parameter
initialised to equilibrium blue phase I structure), but for the entire
simulation system of 128$^3$ lattice sites. (A)  solid volume fraction~1\%
and weak anchoring;  (B) 4\% solid volume fraction and weak anchoring;
(C) 1\% solid volume fraction and strong anchoring; (D) 4\% solid
volume fraction and strong anchoring. The view direction in the
main figure is from the left here. There is a reference
particle at the bottom left in each panel which does not take part
in the simulation.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.35]{support-fig3.pdf}
\end{center}
\caption{As for main manuscript Fig.~1 E--H (initialised via a ``quench''
to generate a disordered network), but for the entire simulation
system of 128$^3$ lattice sites. (E) solid volume fraction 1\% and
weak anchoring; (F) 4\% solid volume fraction and weak anchoring;
(G) 1\% solid volume fraction and strong anchoring; (H) 4\% solid
volume fraction and strong anchoring. Again, there is a reference
particle at the bottom left in each case which does not take part
in the simulation.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.42]{support-fig4.pdf}
\end{center}
\caption{As for main manuscript Fig.~2 A--D (confined geometry with normal
anchoring walls at both top and bottom), but for the entire simulation
system of 256$^2\times$56 lattice sites. Each shows a top view and
side view of the same simulation state. (A) solid volume fraction 1\%
and weak anchoring; (B) 4\% solid volume fraction and weak anchoring;
(C) 1\% solid volume fraction and strong anchoring; (D) 4\% solid
volume fraction and strong anchoring.}
\end{figure}

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.42]{support-fig5.pdf}
\end{center}
\caption{As for main manuscript Fig.~2 E--H (confined geometry with planar
anchoring walls at both top and bottom), but for the entire simulation
system of 256$^2\times$56 lattice sites. Each shows a top view and
side view of the same simulation state. (E) colloid solid volume fraction 1\%
and weak anchoring; (F) 4\% solid volume fraction and weak anchoring;
(G) 1\% solid volume fraction and strong anchoring; (H) 4\% solid
volume fraction and strong anchoring.}
\end{figure}

\begin{figure}[t]
\begin{center}
%%\includegraphics[width=0.95\columnwidth]{s5.jpg}
\end{center}
\caption{Pair correlation function for weak ($WR/K=0.23$) and strong ($WR/K=2.3$) anchoring cases considered in the manuscript. (A,B) BPI initialisation with A) volume fraction $\phi=1${\%} and B) $\phi=5${\%}. (C,D) ``quenched system'' with C) $\phi=1${\%} and D) $\phi=5${\%}.}
\end{figure}

%{\bf Captions for Supplementary Movies}

>>>>

The effect of an electric field can be modelled by including the following
term in the bulk free energy density:
\begin{equation}
-\frac{\epsilon_a}{12\pi} E_{\alpha}Q_{\alpha\beta}E_{\beta},
\end{equation} 
where $E_{\alpha}$ are the components of the electric field, and 
$\epsilon_a$ (here assumed positive) is the dielectric anisotropy of the liquid
crystal. The strength of the electric field is quantified via
one further dimensionless number
\begin{equation}
{\cal E}^2 = \frac{27 \epsilon_a}{32 \pi A_0 \gamma} E_\alpha E_\alpha.
\end{equation}
The quantity ${\cal E}$ will be referred to as the reduced field strength.

{\bf 1.1 Surface free energy}

In addition to the fluid free energy density $f(Q_{\alpha\beta})$, a surface
free energy density is also present to represent the energetic cost of
anchoring at a solid surface. If the preferred order parameter tensor
at the solid surface is $Q^0_{\alpha\beta}$, then the surface free energy
density (per unit area) is
\begin{equation}
f_s(Q_{\alpha\beta}, Q^0_{\alpha\beta})
= {\textstyle \frac{1}{2}}W(Q_{\alpha\beta} - Q^0_{\alpha\beta})^2,
\end{equation}
where $W$ is a constant determining the strength of the anchoring.
The determination of $Q^0_{\alpha\beta}$ for normal and planar anchoring is
discussed below. For a colloid of radius $R$, the strength of the surface
anchoring compared which the bulk fluid elastic constant may be quantified
by the dimensionless parameter $WR/K$. For small values of this parameter,
the presence of a colloid surface should have little impact on the local
fluid LC ordering.

@@

In all simulations the colloidal (hard sphere) radius is $R = 2.3$. To
prevent colloids overlapping at the hard sphere radius, an additional
short range soft potential is included. This takes the form of
$V(h) = \epsilon (\sigma/h)$ where $h$ is the surface to surface separation.
The parameters are $\epsilon = 0.0004$ and $\sigma = 0.1$ in simulation
units. Both potential and resulting force are smoothly matched to zero at
a cut-off distance of $h_c = 0.25$ simulation units.

@@

The parameters for the free energy are chosen to be representative
of blue phase I: chirality $\kappa \sim 0.7$ and reduced temperature
$\tau = -0.25$.
The full free energy parameters are shown in Table~\ref{table:params}.
The order parameter tensor $Q_{\alpha\beta}$ for bulk blue phase I is
initialised from an approximation in high chirality limit
\cite{blue1,oliver1}. Systems of 128$^3$ lattice sites are
used with a pitch length of $p = 32\sqrt{2}$, which accommodates 4$^3$
unit cells with fully periodic boundary conditions.
Simulations at different solid volume fractions (1\% and 4\%), and different 
surface anchoring strengths ($WR/K = 0.23$ and $WR/K = 2.3$) representing
``weak'' and ``strong'' anchoring are run for two million simulation
time steps.
The initial distribution of the colloid positions is set at random
for the required solid volume fraction; the colloids are initially
at rest.

To generate a disordered network of disclination lines, a ``quench''
is performed in which the order parameter is initialised via a
locally chosen random director, and
Eq.~\ref{eq:surf_q} employed with a small amplitude $q^0 = 10^{-7}$ to
provide an order parameter.
We avoid the mean-field spinodal point at $\gamma = 3.0$:
the parameters are the same as for the initial simulations with
pitch $p = 32\sqrt{2}$. The chirality and reduced temperature (see
Table~\ref{table:params}) remain appropriate for BPI.
Colloids are added as before and the simulations run for 2 million
simulation steps.

For simplicity, we do not allow the size of the BPI unit cell to 
dynamically readjust to minimize free energy (there is no 
``redshift''~\cite{blue1}): this would lead to quantitatively different 
values of the free energy, but qualitatively similar structures.

@@

Here, a narrow sandwich of fluid is placed between flat walls in
perpendicular to the narrow coordinate direction, and with periodic
boundaries in the other two directions. The system size used in all
cases is 256$^2 \times$56 lattice sites. Each simulation is a ``quench''
in a similar manner to that used in the bulk: the order parameter is
initialised randomly. The free energy parameters (see Table~\ref{table:params})
are again appropriate for (equilibrium) blue phase I
(chirality $\kappa = 0.8$ and
reduced temperature $\tau = -0.25$). The cholesteric pitch is set
to be $p = 64$, providing a slightly better resolution than the bulk
simulations.

The surface anchoring for colloids is as before: always normal, but with
$WR/K = 0.23$ and $WR/K = 2.3$. The colloid solid volume fractions are
1\% and 4\%. Wall anchoring is either normal or planar with $W/K = 1$
in simulation units. All simulations are run for two million simulation
time steps.

@@

For the simulations of BPIII with colloidal particles a system size of
$128^3$ is used with a cholesteric pitch of $p=32$ in lattice units. 
The simulations are
performed at chirality $\kappa=2.6$ and reduced temperature $\tau=-0.25$.  
An initial configuration of
randomly oriented and positioned double twist cylinders is used,
from which an amorphous BPIII network emerges \cite{oliver-bp3}. This structure
is equilibrated for $6\times10^{5}$ LB time steps until no significant
further evolution is observed.

The colloidal particles with weak
normal anchoring ($WR / K= 0.23$), are then inserted with randomly chosen positions,
and the composite system is equilibrated for another $1.2\times10^{6}$
time steps.
At the end of the equilibration phase the uniform electric field
with reduced field strength ${\cal E} = 0.8$ 
is switched on (oriented along the $z$-direction), and the simulation
run for a further $1.2\times10^{6}$ time steps.
This is found to be sufficiently long for a hexagonal structure to
emerge and saturate.
If the electric field is then switched off, and the system quickly relaxes
to a metastable state which shows a residual anisotropy in the
colloidal distribution. The total length of this final relaxation phase
is $1\times10^6$ LB time steps.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
         & $A_0$ & $\gamma$ & $K$ & $p$ & $\kappa$ & $\tau$ & $WR/K$\\
\hline 
Bulk BPI & 0.01 & 3.086 & 0.007061 & 32$\sqrt{2}$ & 0.6902 & -0.2500 &
0.23, 2.3\\
\hline
Bulk Quench & 0.01 & 3.086 & 0.007061 & 32$\sqrt{2}$ & 0.6902 & -0.2500 &
0.23, 2.3 \\
\hline
Confined & 0.01 & 3.086 & 0.01897 & 64 & 0.8 & -0.2500 & 0.23, 2.3\\
\hline
External field & 0.004& 3.086 & 0.02 & 32 & 2.6 & -0.2500 & 0.23\\ 
\hline
\end{tabular}
\end{center}
\caption{Free energy parameters used in the simulations. Bulk BPI
correspond to Fig.~1 (A-D), bulk quench to Fig.~1 (E--H); confined
geometries are shown in Fig.~2; and the external field case is
relevant for Fig.~3.}
\label{table:params}
\end{table}

{\bf 4.4 Disclination rendering}

@@

To get from simulation to physical units, a calibration of scales of
length, energy and time in the simulations is required. For similar
mappings see,
e.g.~\cite{denniston2}.

The length scale can be set by mapping the half pitch to a realistic blue
phase unit cell size, which is in the 100--500~nm range \cite{blue1}. 
For bulk simulations (Fig.~1 and Fig.~3 in the main text), a suitable choice
is one where one simulation unit (lattice site) corresponds to 10~nm. Therefore
the colloidal size in Figs.~1 and~3 corresponds to $\sim$ 50~nm.

To obtain an energy scale, it is possible to choose $A_0 \simeq 10^6$~Pa,
which is reasonable following Ref.~\cite{blue1} (this is also the choice in
Ref.~\cite{oliver2}). This choice leads to a simulation unit of
free energy (or stress) equal to $10^{8}$~Pa in Figs.~1 and~3.
From the energy and length scales one can map elastic constants to
physical units: for instance, the simulation in Fig.~1 corresponds to
a liquid crystal with splay, bend and twist (Frank) elastic constants
equal to 35 pN (or equivalently $K=70$~pN).

The timescale calibration may be obtained from (see e.g.~\cite{denniston})
\begin{equation}
\gamma_1=\frac{2q^2}{\Gamma},
\end{equation}
which relates the rotational viscosity $\gamma_1$ to the ordering strength
$q$ and the order parameter mobility $\Gamma$.
In the present simulations, $\Gamma = 0.5$ and the value of $\gamma$
for the simulation in Fig.~1 leads to $q=1/2$.   
For real liquid crystalline materials, $\gamma_1$ usually lies in range
$10^{-2}-1$ Pa s~\cite{deGennes}; for definiteness say
$\gamma_1 = 1$~Pa~s (equivalently 10 poise). Given the previous mapping for
free energy (or stress) units, this leads to a simulation unit of time equal
to $10^{-8}$~s. The total simulated time in Figs. 1-3 is therefore in the
millisecond range.

To calibrate the electric field strength 
one can equate the value of the dimensionless parameter ${\cal E}$ in the 
simulations with that of a hypothetical experiment.
Given the previous value for $A_0$, and a dielectric anisotropy of 20
(or equivalently $\epsilon_a=20\epsilon_0$, with $\epsilon_0$ the
dielectric permittivity of vacuum), an electric field
of 100 V/$\mu$m is obtained. This corresponds to a value of
${\cal E}\sim 0.4$ (considering a value of $\gamma=3.0857$ as in
Fig.~3 of the main text).

Given these mappings, all quantities can be easily transformed from
simulation to physical units and vice versa. For example Figs.~1 and~2
use an isotropic viscosity $\eta \simeq$~0.01 and~0.1 in simulation units,
which translate into $0.01-0.1$~Pa~s respectively. This is sensible (if 
slightly low) for a molecular nematogen in the isotropic phase. 
The effective viscosity in ordered phases is higher, but this is ensured
in our model by the coupling to the order parameter.
Note that the density is set to unity in the simulations: this
corresponds to a fluid density which is about a thousand times larger
than in experiments. This makes no difference in practice as
the Reynolds number $= \rho V\Lambda/\eta$ (where $V$ is a typical
velocity, in our simulations around $10^{-5}$ in lattice units, 
and $\Lambda$ is a typical length scale, e.g. the size of a unit cell) 
remains small enough~\cite{codef}.

To summarize, the simulations represent BP-forming materials, with the
interpretation of the simulation units for length, time, and energy
density being close to 10~nm, 10~ns, and 100~MPa respectively.

@@

\bibitem{blue1}
D.C. Wright and N.D. Mermin,
%  Crystalline liquids -- the blue phases.
{\it Rev. Mod. Phys.} {\bf 61,} 385-432 (1989).
\bibitem{deGennes}
P.-G. de Gennes, J. Prost,
{\it The Physics of Liquid Crystals} (Oxford Univ. Press, New York, 1993).
%\bibitem{marendu1}
%D. Marenduzzo, E.  Orlandini, M. E. Cates, J. M. Yeomans,
%{\rm Steady-state hydrodynamic instabilities of active liquid crystals:
%Hybrid lattice Boltzmann simulations.}
%{\it Phys. Rev. E} {\bf 76}, 031921 (2007).
%\bibitem{marendu2}
%O. Henrich, D. Marenduzzo, K. Stratford, M. E. Cates, 
%{\rm Domain growth in cholesteric blue phases:
%hybrid lattice Boltzmann simulations}.
%{\it Comput. Math. Appl.}, doi:10.1016/j.camwa.2009.08.047[dx.doi.org] (2009).
%\bibitem{marendu3}
%M. E. Cates, O. Henrich, D. Marenduzzo, K. Stratford,
%{\rm  Lattice Boltzmann simulations of liquid crystalline fluids:
%active gels and blue phases.}
%{\it Soft Matter} {\bf 5}, 3791-3800 (2009).
\bibitem{oliver1}
O. Henrich, D. Marenduzzo, K. Stratford, and M.E. Cates,
%Thermodynamics of blue phases in electric fields
\textit{Phys. Rev. E}, \textbf{81} 031706 (2010).
\bibitem{beris}
A. N. Beris, B. J. Edwards, 
{\it Thermodynamics of Flowing Systems with Internal Microstructure}
(Oxford Univ. Press, New York, 1994).
\bibitem{ladd94}
A.J.C. Ladd,
\textit{J. Fluid Mech.} \textbf{271}, 285 (1994); \textit{ibid.} \textbf{271}
311 (1994).
\bibitem{nguyen2002}
N.-Q. Nguyen and A.J.C. Ladd,
\textit{Phys. Rev. E} \textbf{66} 046708 (2002).
\bibitem{juho}
J.S. Lintuvuori, D. Marenduzzo, K. Stratford, and M.E. Cates,
\textit{J. Mater. Chem.} \textbf{20}, 10547 (2010).
\bibitem{denniston}
C. Denniston, E. Orlandini, J. M. Yeomans, 
\textit{Phys. Rev. E} \textbf{63}, 056702 (2001).
\bibitem{fournier2005}
J.-B. Fournier and P. Galatola,
\textit{Europhys. Lett.} \textbf{72}, 403 (2005).
\bibitem{oliver-bp3}
O. Henrich, K. Stratford, M.~E. Cates, D. Marenduzzo,
{\it Phys. Rev. Lett.} {\bf 106}, 107802 (2011).
\bibitem{denniston2}
C. Denniston, D. Marenduzzo, E. Orlandini, J. M.  Yeomans, 
\textit{Philos. Trans. R. Soc. London, Ser. A} \textbf{362}, 1745 (2004).
\bibitem{oliver2}
O. Henrich, K. Stratford, D. Marenduzzo, M. E. Cates, 
\textit{Proc. Acad. Sci. USA} \textbf{107}, 1312 (2010).
\bibitem{codef} M. E. Cates,  et al., 
\textit{J. Phys. Condens. Matter} \textbf{16},  3903 (2004).

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[scale=0.8]{support-fig1.png}
\end{center}
\caption{
(A) Schematic representation of the director field
in a cholesteric (left) and within a double twist cylinder (right).
(B) Snapshot of the disclination network in an equilibrated 
blue phase I structure. (C) Same as (B) for an amorphous blue phase III
network.}
\end{figure}

\newpage

\begin{figure}[!h]
\begin{center}
\includegraphics[scale=0.35]{support-fig2.pdf}
\end{center}
\caption{As for main manuscript Fig.~1 A--D
(with liquid crystal order parameter
initialised to equilibrium blue phase I structure), but for the entire
simulation system of 128$^3$ lattice sites. (A)  solid volume fraction~1\%
and weak anchoring;  (B) 4\% solid volume fraction and weak anchoring;
(C) 1\% solid volume fraction and strong anchoring; (D) 4\% solid
volume fraction and strong anchoring. The view direction in the
main figure is from the left here. There is a reference
particle at the bottom left in each panel which does not take part
in the simulation.}
\end{figure}

\newpage

\begin{figure}[!h]
\begin{center}
\includegraphics[scale=0.35]{support-fig3.pdf}
\end{center}
\caption{As for main manuscript Fig.~1 E--H (initialised via a ``quench''
to generate a disordered network), but for the entire simulation
system of 128$^3$ lattice sites. (E) solid volume fraction 1\% and
weak anchoring; (F) 4\% solid volume fraction and weak anchoring;
(G) 1\% solid volume fraction and strong anchoring; (H) 4\% solid
volume fraction and strong anchoring. Again, there is a reference
particle at the bottom left in each case which does not take part
in the simulation.}
\end{figure}

\newpage

\begin{figure}[!h]
\begin{center}
\includegraphics[scale=0.49]{support-fig4.png}
\end{center}
\caption{Plot of the free energy (per BP unit cell) versus time
for the simulations corresponding to Fig. 1A-H in the main text
(the labels correspond to those in Fig. 1).
}
\end{figure}

\newpage

\begin{figure}[!h]
\begin{center}
\includegraphics[scale=0.42]{support-fig5.pdf}
\end{center}
\caption{As for main manuscript Fig.~2 A--D (confined geometry with normal
anchoring walls at both top and bottom), but for the entire simulation
system of 256$^2\times$56 lattice sites. Each shows a top view and
side view of the same simulation state. (A) solid volume fraction 1\%
and weak anchoring; (B) 4\% solid volume fraction and weak anchoring;
(C) 1\% solid volume fraction and strong anchoring; (D) 4\% solid
volume fraction and strong anchoring.}
\end{figure}

\newpage

\begin{figure}[!h]
\begin{center}
\includegraphics[scale=0.42]{support-fig6.pdf}
\end{center}
\caption{As for main manuscript Fig.~2 E--H (confined geometry with planar
anchoring walls at both top and bottom), but for the entire simulation
system of 256$^2\times$56 lattice sites. Each shows a top view and
side view of the same simulation state. (E) colloid solid volume fraction 1\%
and weak anchoring; (F) 4\% solid volume fraction and weak anchoring;
(G) 1\% solid volume fraction and strong anchoring; (H) 4\% solid
volume fraction and strong anchoring.}
\end{figure}

\newpage

\begin{figure}[!h]
\begin{center}
\includegraphics[width=0.95\columnwidth]{support-fig7.pdf}
\end{center}
\caption{Panels A and B as for Fig. 3, but with the disclination lines
removed for clarity. Panel C shows the situation when the field is
switched off: a residual anisotropic structure remains which reflects the
structure induced by the electric field. The electric field direction
when switched on (panel B) is into the plane of the paper.}
\end{figure}

%\newpage

{\bf Captions for Supplementary Movies} \\

{\bf Supplementary Movie 1:} 
The final state of the simulation as shown in Fig.~1A. The movie
rotates the configuration to reveal more clearly its structure.\\

{\bf Supplementary Movie 2:} 
The final state of the simulation as shown in Fig.~1B. The movie
rotates the configuration to reveal more clearly its structure.\\

{\bf Supplementary Movie 3:} 
The final state of the simulation as shown in Fig.~1C. The movie
rotates the configuration to reveal more clearly its structure.\\

{\bf Supplementary Movie 4:} 
The final state of the simulation as shown in Fig.~1D. The movie
rotates the configuration to reveal more clearly its structure.\\

{\bf Supplementary Movie 5:} 
The configuration after the initial relaxation phase in Fig.~3. The movie
shows a rotation of the configuration. The distribution of the colloidal
particles is isotropic, with the particles forming clusters and having a
tendency to occupy positions at the intersection of the BPIII disclination
lines.\\

{\bf Supplementary Movie 6:} 
Configuration in a homogeneous electric field with strength
corresponding to ${\cal E}=0.8$. The movie
shows a rotation of the configuration.
A hexagonal pattern emerges where the particles form chains
situated at the vertices of the honeycomb-like structure. \\
